\ttitle

### 2015-07-12 Chapters/Introduction.tex
Minor changes to Introduction
- Include section labels for POPF, PPH
- vfill/hfill in ISGPS tables

\subsubsection{Postoperative pancreatic fistula}
Postoperative pancreatic fistula is one of the most dreaded complications after a pancreaticoduodenectomy and can be associated with significant short-term morbidity as well as long-term disability. The reported incidence of postoperative pancreatic fistula varies from 2\% to 30\% after pancreaticoduodenectomy.\parencite{yeo_six_1997,deoliveira_assessment_2006,bassi_postoperative_2005,winter_biochemical_2007,pratt_risk_2008} 
The variation in reported incidence has been largely due to lack of clear definition of what constituted a postoperative pancreatic fistula. It can be a result of breakdown or poor healing at the pancreaticojejunostomy/pancreaticogastrostomy or may be the result of direct parenchymal leak unrelated to the anastomosis. It is now generally accepted that 1 in 4 patients will develop a pancreatic fistula as defined by the International Study Group for Pancreatic Fistula (ISGPF) which has published a consensus statement on the definition and grading of postoperative pancreatic fistula.\parencite{bassi_postoperative_2005} A postoperative pancreatic fistula is defined as drain output of any measurable quantity after the third postoperative day with amylase content greater than three times the upper limit of the normal serum amylase value at the laboratory used for testing. Three grades of postoperative pancreatic fistula have been defined based on clinical severity as described in Table \ref{table:isgps_popf} on p\pageref{table:isgps_popf}. Grade B and C fistulae are considered to be clinically significant in that they alter patient management and are often associated with other secondary complications such as intra-abdominal sepsis, post-pancreatectomy haemorrhage, delayed gastric emptying as well as need for intervention (either radiological or operative) and/or prolonged critical care support.

\subsubsection{Post-pancreatectomy haemorrhage}
Post-pancreatectomy haemorrhage is reported to occur in 1 to 8\% of patients undergoing pancreaticoduodenectomy. However, it accounts for 11\% to 38\% of mortality after pancreaticoduodenectomy. Post-pancreatectomy haemorrhage may either be intra-luminal into the gastrointestinal tract or intra-abdominal into the peritoneal/retro-peritoneal space. Post-pancreatectomy haemorrhage may be from any of a number of potential sources although bleeding from the stump of the gastroduodenal artery is the most common cause. Other potential sources include suture lines at the anastomoses, gastric/duodenal ulcers or diffuse gastritis, pseudoaneurysms of the gastro-duodenal, splenic or rarely the hepatic artery or rarely, haemobilia.

>>>>

\subsubsection{Postoperative pancreatic fistula}
\label{sec:ch_intro_POPF}
Postoperative pancreatic fistula is one of the most dreaded complications after a pancreaticoduodenectomy and can be associated with significant short-term morbidity as well as long-term disability. The reported incidence of postoperative pancreatic fistula varies from 2\% to 30\% after pancreaticoduodenectomy.\parencite{yeo_six_1997,deoliveira_assessment_2006,bassi_postoperative_2005,winter_biochemical_2007,pratt_risk_2008} 
The variation in reported incidence has been largely due to lack of clear definition of what constituted a postoperative pancreatic fistula. It can be a result of breakdown or poor healing at the pancreaticojejunostomy/pancreaticogastrostomy or may be the result of direct parenchymal leak unrelated to the anastomosis. It is now generally accepted that 1 in 4 patients will develop a pancreatic fistula as defined by the International Study Group for Pancreatic Fistula (ISGPF) which has published a consensus statement on the definition and grading of postoperative pancreatic fistula.\parencite{bassi_postoperative_2005} A postoperative pancreatic fistula is defined as drain output of any measurable quantity after the third postoperative day with amylase content greater than three times the upper limit of the normal serum amylase value at the laboratory used for testing. Three grades of postoperative pancreatic fistula have been defined based on clinical severity as described in Table \ref{table:isgps_popf} on p\pageref{table:isgps_popf}. Grade B and C fistulae are considered to be clinically significant in that they alter patient management and are often associated with other secondary complications such as intra-abdominal sepsis, post-pancreatectomy haemorrhage, delayed gastric emptying as well as need for intervention (either radiological or operative) and/or prolonged critical care support.

\subsubsection{Post-pancreatectomy haemorrhage}
\label{sec:ch_intro_PPH}
Post-pancreatectomy haemorrhage is reported to occur in 1 to 8\% of patients undergoing pancreaticoduodenectomy. However, it accounts for 11\% to 38\% of mortality after pancreaticoduodenectomy. Post-pancreatectomy haemorrhage may either be intra-luminal into the gastrointestinal tract or intra-abdominal into the peritoneal/retro-peritoneal space. Post-pancreatectomy haemorrhage may be from any of a number of potential sources although bleeding from the stump of the gastroduodenal artery is the most common cause. Other potential sources include suture lines at the anastomoses, gastric/duodenal ulcers or diffuse gastritis, pseudoaneurysms of the gastro-duodenal, splenic or rarely the hepatic artery or rarely, haemobilia.
